MinCall --- MinION end2end deep learning basecaller

### 2017-06-30 paper/main.tex
Added initial results CIGAR section

@@

>>>>

Developed tool was compared with other available basecallers that support R9 chemistry. This includes third-party basically DeepNano and official basecallers by Oxford Nanopore (cloud-based Metrichor and Nanonet).
The fact that ground truth is not known makes evaluation difficult.
Different methods for evaluation were used to get clearer information about each basecaller.

A portion of the read length that aligns as correctly is called match\_rate. Same goes for mismatches and insertions.
Sum of all matches, mismatches, and insertions is equal to the reads length \ref{eq:read_len}. For specific details see appendix~\ref{app:evaluation:CIGAR}. Results on E.Coli test set with Graphmap aligner are shown in table~\ref{tbl:ecoli_rates}. Evaluation with BWM-MEM and on lambda can be found in appendix~\ref{app:evaluation:CIGAR} under tables~\ref{tbl:ecoli_rates_bwa},~\ref{tbl:lambda_rates} and~\ref{tbl:lambda_rates_bwa}. Furthermore we plot Kernel Density estimation(KDE) plots for each mentioned statistic on E.Coli dataset in figure~\ref{fg:ecoli_kde}.

\begin{table}[]
	\caption{Alignment specifications of Ecoli R9 basecalled reads using GraphMap}
	\label{tbl:ecoli_rates}
	\centering
	\begin{tabular}{lcccc}
		\toprule
		{} &  \thead{Match \% \\(median)} &  \thead{Mismatch \% \\(median)} &  \thead{Insertion \% \\(median)} &  \thead{Deletion \% \\(median)} \\
		\midrule
		DeepNano   &                  90.254762 &                      6.452852 &                       \textbf{3.274420} &                     11.829965 \\
		Metrichor  &                  90.560455 &                      5.688105 &                       3.660381 &                      8.328271 \\
		Nanonet    &                  90.607674 &                      5.608912 &                       3.652791 &                      8.299046 \\
		MinCall    &                  \textbf{91.408591} &                     \textbf{ 5.019141} &                       3.477739 &                      \textbf{7.471608 }\\
		\bottomrule
	\end{tabular}
\end{table}

\begin{figure}[!htb]
	\begin{center}
		\includegraphics[width=0.8\textwidth]{./imgs/results/ecoli/kde_cigar.png}
		\caption{KDE plot for distribution of percentage of alimnment operations for ecoli}
		\label{fg:ecoli_kde}
	\end{center}
\end{figure}

@@

\section{Evaluation metrics}

\subsection{CIGAR derived metrics}
\label{app:evaluation:CIGAR}

A portion of the read length that aligns as correctly is called match\_rate. Same goes for mismatches and insertions.
Sum of all matches, mismatches, and insertions is equal to the reads length \ref{eq:read_len}.
Here is the specific equation list:

\begin{equation}
\begin{gathered}
\label{eq:read_len}
read\_len =  n\_matches + n\_mismatches + n\_insertions
\end{gathered}
\end{equation}
\begin{equation}
\begin{gathered}
match\_rate = \dfrac{n\_matches}{read\_length}
\end{gathered}
\end{equation}
\begin{equation}
\begin{gathered}
missmatch\_rate = \dfrac{n\_mismatches}{read\_length}
\end{gathered}
\end{equation}
\begin{equation}
\begin{gathered}
insertion\_rate = \dfrac{n\_insertions}{read\_length}
\end{gathered}
\end{equation}

\begin{equation}
\begin{gathered}
match\_rate  + snp\_rate + insertion\_rate = 1
\end{gathered}
\end{equation}

Deletion rate is defined as a total number of deletions in the alignment over the length of the aligned read.

\begin{equation}
\begin{gathered}
deletion\_rate = \dfrac{n\_deletion}{read\_length}
\end{gathered}
\end{equation}

\begin{table}[]
	\caption{Alignment specifications of Ecoli R9 basecalled reads using BWA mem}
	\label{tbl:ecoli_rates_bwa}
	\centering
	\begin{tabular}{lcccc}
		\toprule
		{} &  \thead{Match \% \\(median)} &  \thead{Mismatch \% \\(median)} &  \thead{Insertion \% \\(median)} &  \thead{Deletion \% \\(median)} \\
		\midrule
		DeepNano   &                  90.254762 &                      6.452852 &                       3.274420 &                     11.829965 \\
		Metrichor  &                  90.595441 &                      6.869543 &                       2.531646 &                      7.567381 \\
		Nanonet    &                  90.988989 &                      6.674760 &                     \textbf{  2.348552} &                      7.698530 \\
		MinCall    &                 \textbf{ 91.470588} &                      \textbf{5.929204 }&                       2.477283 &                     \textbf{ 6.970362} \\
		\bottomrule
	\end{tabular}

\end{table}
\begin{table}[]
	\caption{Alignment specifications of Lambda basecalled reads using GraphMap}
	\label{tbl:lambda_rates}
	\centering
	\begin{tabular}{lcccc}
		\toprule
		{} &  \thead{Match \% \\(median)} &  \thead{Mismatch \% \\(median)} &  \thead{Insertion \% \\(median)} &  \thead{Deletion \% \\(median)} \\
		\midrule
		DeepNano   &                  86.997687 &                      9.623494 &                       3.442490 &                     16.052830 \\
		Metrichor  &                  87.714988 &                      7.835052 &                       4.093851 &                     \textbf{10.757491} \\
		Nanonet    &                  88.415611 &                      8.178372 &                       3.629653 &                     11.793022 \\
		MinCall    &                 \textbf{ 89.694482} &                     \textbf{ 7.238095} &                       \textbf{3.078796} &                     13.450292 \\
		\bottomrule
	\end{tabular}

\end{table}
\begin{table}[]
	\caption{Alignment specifications of lambda R9 basecalled reads using BWA mem}
	\label{tbl:lambda_rates_bwa}
	\centering
	\begin{tabular}{lcccc}
		\toprule
		{} &  \thead{Match \% \\(median)} &  \thead{Mismatch \% \\(median)} &  \thead{Insertion \% \\(median)} &  \thead{Deletion \% \\(median)} \\
		\midrule

		DeepNano   &                  86.625973 &                     11.288361 &                       2.098225 &                     14.648308 \\
		Metrichor  &                  87.294093 &                     10.109186 &                       2.376476 &                      \textbf{9.645323 }\\
		Nanonet    &                  87.767037 &                     10.017598 &                       2.354248 &                     10.597232 \\
		MinCall    &                 \textbf{ 89.049870} &                     \textbf{ 9.480883} &                       \textbf{1.615188 }&                     12.962441 \\
		\bottomrule
	\end{tabular}
\end{table}
